The Processing of American Indian/Alaska Native Children through non-Tribal

### 2019-12-19 ms.rmd
ff

```{r include=F}
rm(list=ls()); gc()
library(tidyverse)
source("lifetable.r")

pop<-read_csv("./data/pop_aian_nat.csv")
afcars<-read_csv("./data/afcars_subset.csv",
                 col_types = cols(stfcid="c"))

afcars_all<-read_csv("./data/afcars_id_xwalk.csv",
                     col_types = cols(stfcid="c"))

ncands<-read_csv("./data/ncands_subset.csv")

xwalk<-ncands_xwalk %>% 
  left_join(afcars_all) %>% 
  filter(!(is.na(fy))) %>% 
  select(stfcid, rptdt, fy) %>% 
  rename(fc_year = fy)

@@

In contrast to per capita event rates, life table estimates allow us to simulate the incidence of an event over the course of a child's life. These estimates adjust for variable age-specific risk to provide a more comprehensive single estimate of event incidence in the population.

@@

$P(\textsf{investigation})$

Establish the probability that a child was ever investigated.

```{r ncandstables.pop}
ncands_first<-ncands %>% 
  select(.imp, chid, chage, rptdt) %>% 
  group_by(chid) %>% 
  filter(rptdt==min(rptdt)) %>% 
  distinct() %>% 
  ungroup() %>% 
  mutate(year = year(rptdt))

ncands_first_nat<-ncands_first %>% 
  group_by(chage, year) %>% 
  summarise(var = n()) %>% 
  rename(age = chage) %>% 
  left_join(pop) %>% 
  ungroup()

investigation_tables<-list()
years<-unique(ncands_first_nat$year)
for(i in 1:length(years)){
  investigation_tables[[i]]<-make_life_table(ncands_first_nat %>% 
                  filter(year==years[i]))
}

@@

@@

```{r fctables.ncands}
ncands_first<-ncands %>% 
  filter(rptvictim==1) %>% 
  group_by(chid) %>% 
  filter(rptdt==min(rptdt))

ncands_first_subst_nat<-ncands_first %>% 
  group_by(chage, year) %>% 
  summarise(var = n()) %>% 
  rename(age = chage) %>% 
  left_join(pop) %>% 
  ungroup()

@@

```{r fctables.pop}
### look at first fc placement
afcars_first <- afcars %>% 
  group_by(stfcid) %>% 
  filter(age == min(age) & year == min(year)) %>% 
  select(stfcid, age, year) %>% 
  distinct()
### this matches length(unique(afcars$stfcid))

afcars_first_nat<-afcars_first %>% 
  group_by(age, year) %>% 
  summarise(var = n()) %>% 
  left_join(pop) %>% 
  ungroup()

@@

## $P(FC)$ using NCANDS

```{r}
### look at first fc placement
ncands_first_fc <- ncands %>% 
  filter(!(is.na(afcarsid))) %>% 
  group_by(chid) %>% 
  filter(rptdt==min(rptdt)) %>% 
  select(chid, afcarsid, chage, year) %>% 
  ungroup() %>% 
  distinct()
### this matches length(unique(afcars$stfcid))

ncands_fc_nat<-ncands_first_fc %>% 
  rename(age = chage) %>% 
  group_by(age, year) %>% 
  summarise(var = n()) %>% 
  left_join(pop) %>% 
  ungroup()

## $P(entered, investigation)$ using ncands
```{r fctables.ncands}
ncands_fc_first<-ncands %>% 
  filter(ever_entered==T) %>% 
  group_by(stfcid) %>% 
  filter(chage==min(chage)) %>% 
  ungroup() %>% 
  select(stfcid, year, chage) %>% 
  distinct()

ncands_fc_nat<-ncands_fc_first %>% 
  group_by(year, chage) %>% 
  summarise(var = n()) %>% 
  rename(age = chage) %>% 
  left_join(pop) %>% 
  ungroup()

### run lifetable by year
ncands_fc_tables<-list()
years<-unique(ncands_fc_nat$year)
for(i in 1:length(years)){
  ncands_fc_tables[[i]]<-make_life_table(ncands_fc_nat %>% 
                  filter(year==years[i]))
}

ncands_fc_tables<-bind_rows(ncands_fc_tables)
```

@@

afcars_inv<-afcars_first %>% 
  left_join(ncands_xwalk %>% 
              select(stfcid, rptdt) %>% 
              mutate(ncands_inv = T))

afcars_inv<-afcars_inv %>% 
  ungroup() %>% 
  mutate(rpt_year = year(rptdt)) %>% 
  filter(!(is.na(ncands_inv)))

## remove when year(rptdt)>year

afcars_inv<-afcars_inv %>% 
  filter(year>=rpt_year)

>>>>

```{r include=F}
rm(list=ls()); gc()
library(tidyverse)
library(lubridate)
source("lifetable.r")

@@

In contrast to per capita event rates, life table estimates allow us to simulate the incidence of an event over the course of a child's life. These estimates adjust for variable age-specific risk and population size to provide a more comprehensive single estimate of event incidence in the population.

@@

```{r ncandstables.pop}
### read in new first investigation csv
### run lifetable by year

investigation_tables<-list()
years<-unique(ncands_first_nat$year)
race_ethn<-unique(ncands_first$race_ethn)
for(i in 1:length(years)){
  for(r in 1:length(race_ethn)){
  investigation_tables[[i]]<-make_life_table(ncands_first_nat %>% 
                  filter(year==years[i]) %>% 
                    filter(race_ethn == race_ethn[r]))
  }
}

@@

FIG 1: TS plot of age 17 cumulative investigation risk by year

@@

```{r fctables.ncands}

@@

```{r fctables.pop}

@@

@@

### lifetable for p(inv,entered)
